---
title: "Pruebas Saber 3, 5 y 9 - Región Pacífico"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll #scroll #fill -fixed
    theme: sandstone ## broad customization 
    #css: cssfile.css ## more layout control 
    scroll_page: True
  runtime: shiny 
---

```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(tidyverse)
require(here)
require(sf)
require(ggplot2)

require(osmdata)
require(tigris)
require(viridis)
require(tmap)
require(elevatr)
library(gghighlight)
library(geomtextpath)
```

```{r}
### Load data
df <- read_csv(here('Data/df_ETCs_ALL.csv'))
#head(df)

### Specify filtered units (ETCs)
etcs_filtradas <- c('Popayán', 'Pasto','Ipiales','Tumaco','Quibdó','Palmira','Yumbo',
                      'Tuluá','Cali','Cartago','Buenaventura', 'Buga', 'Jamundí',
                      'Chocó', 'Valle del Cauca', 'Cauca','Nariño')
etcs_filtradas <-sort(etcs_filtradas)

df_filered <- df[df$ENTIDAD %in% etcs_filtradas,]

### set grados, unique 
grados <- unique(df_filered$grado)

### Set areas, unique
areas <- unique(df_filered$area)


```

Context
=======================================================================

Sidebar {.sidebar}
-----------------------------------------------------------------------

**Context of the evaluations**
<br>
The Pruebas Saber 3º, 5º y 9º are evaluations as part a strategy of the Colombian national government, through the Colombian Institute for the Evaluation of Education (ICFES), to measure the learning processes of students in the country. The other main purposes of this effort are: <br>
- to evaluate different areas of knowledge; <br>
- to understand the factors associated with learning development; and <br>
- to measure the socio-emotional skills of students.
<br>
<br>
**Characteristics of the evaluations** <br>
These evaluations are a **sampled** and not _census_ - that is, not all educational institutions participate in the evaluations, but the results seek to be representative of the different territorial entities. The tests were applied up to 2017 and were reviewed for several years, giving way to a second cycle starting in **2022**. The scores from these two stages are not comparable: the most recent results are the new **baseline** for measuring competencies from now on. To learn more details about the tests and their application, visit the following website: 


Row {data-width=100}
-----------------------------------------------------------------------
### **Purpose of the dashboard**
This dashboard has been created with the purpose of visualizing the results of the Saber 3, 5, and 9 evaluations in the Pacific region of Colombia. It is not only about taking advantage of microdata at the level of Certified Territorial Entities (ETCs), but also about providing **actionable information** for decision makers in the territory and thus allowing better responses to the needs of each educational context.
<br>
**Content**
<br>
- On this first page, you will find a map of the ETCs in the region and the number of evaluations that were carried out per year. <br>
- On the second page, you will find the average scores for each ETC in the years when evaluations were conducted. There are also visualizations of the percentages by performance levels for each of the evaluated areas. <br>
- Finally, the last page allows a comparison of the average scores of the different ETCs in the region for each grade and subject area.



Column {data-height=500}
-----------------------------------------------------------------------
### **Map of the Pacific Region** {data-padding=0}

```{r}
### Load shapefiles for Colombia
colombia0  <- sf::st_read(here('Data/gadm41_COL_shp/gadm41_COL_0.shp'))
colombia1  <- sf::st_read(here('Data/gadm41_COL_shp/gadm41_COL_1.shp'))
colombia2  <- sf::st_read(here('Data/gadm41_COL_shp/gadm41_COL_2.shp'))

### Specify departments in the Pacific region
departamentos_pacifico <- c('Chocó', 'Valle del Cauca', 'Cauca','Nariño')
pacifico <- colombia1[colombia1$NAME_1 %in% departamentos_pacifico ,]

### Specify cities within regions
ciudades <- c('Popayán', 'Pasto','Ipiales','Tumaco','Quibdó','Palmira','Yumbo',
 'Tuluá','Cali','Cartago','Buenaventura', 'Buga', 'Jamundí')
ciudades_map <- colombia2[colombia2$NAME_2 %in% ciudades ,]


###FALTAN "Pasto" "Cali"  "Buga" 
#ciudades[!ciudades %in% ciudades_map$ENTIDAD]

###Merge departments and cities
pacifico <- pacifico %>% 
  select("COUNTRY", "NAME_1", "TYPE_1", "geometry") %>%
  rename("TYPE"="TYPE_1",
         "Dept_name"="NAME_1")
pacifico$ENTIDAD <-pacifico$Dept_name

ciudades_map <- ciudades_map %>% 
  select("COUNTRY", "NAME_1", "NAME_2", "TYPE_2", "geometry") %>%
  rename("TYPE"="TYPE_2",
         "Dept_name"="NAME_1",
         "ENTIDAD"="NAME_2")
mapa_pacifico_todo <- rbind(pacifico, ciudades_map)

### COMPUTE number of evaluations
evals_por_entidad <- df_filered %>%
  group_by(ENTIDAD, ano) %>%
  summarize(suma_N = sum(N))

evals_entidad_long <- pivot_wider(evals_por_entidad, names_from=ano, values_from=suma_N)

### Merge mapa con eval
mapa_pacifico_todo_info <- merge(mapa_pacifico_todo, evals_entidad_long)

#### Zoom in a bit??
#bb <- osmdata::getbb(place_name = 'Colombia')
#bb

# ### Plot map
# ggplot() +
#   #geom_sf(data = colombia0, color = "gray60") +
#   geom_sf(data = pacifico, fill="#518FB5") +
#   geom_sf(data = ciudades_map, fill="#e34a33") +
#   #coord_sf(xlim = bb[1, ], ylim = bb[2, ])
#   theme_void()

interactive_map <- tm_shape(mapa_pacifico_todo_info) +
  tm_polygons("TYPE",
              palette = c("#518FB5","#e34a33"),
              border.col = "gray90",
              lwd = 0.1,
              popup.vars = c("2009","2012","2013","2014","2015","2016","2017")) +
  tm_layout(legend.outside = TRUE)
interactive_map
tmap_mode("view")

# tmap_leaflet(
#       interactive_map,
#       mode = "view",
#       height = 500
#     )
```

```{r eval=FALSE}
tmap_mode("view")

tm_shape(mapa_pacifico_todo_info) +
  tm_polygons("TYPE",
              palette = c("#518FB5","#e34a33"),
              border.col = "gray90",
              lwd = 0.1,
              popup.vars = c("suma_N")) +
  tm_layout(legend.outside = TRUE)

```

### **Map of Colombia** {data-padding=0}
```{r}
# ### Load shapefiles for Colombia
# colombia0  <- sf::st_read(here('Data/gadm41_COL_shp/gadm41_COL_0.shp'))
# colombia1  <- sf::st_read(here('Data/gadm41_COL_shp/gadm41_COL_1.shp'))
# colombia2  <- sf::st_read(here('Data/gadm41_COL_shp/gadm41_COL_2.shp'))
# 
# ### Specify departments in the Pacific region
# departamentos_pacifico <- c('Chocó', 'Valle del Cauca', 'Cauca','Nariño')
# pacifico <- colombia1[colombia1$NAME_1 %in% departamentos_pacifico ,]
# 
# ### Specify cities within regions
# ciudades <- c('Popayán', 'Pasto','Ipiales','Tumaco','Quibdó','Palmira','Yumbo',
#  'Tuluá','Cali','Cartago','Buenaventura', 'Buga', 'Jamundí')
# ciudades_map <- colombia2[colombia2$NAME_2 %in% ciudades ,]
# 
# #### Zoom in a bit??
# #bb <- osmdata::getbb(place_name = 'Colombia')
# #bb

### Plot map
ggplot() +
  geom_sf(data = colombia0, color = "gray60") +
  geom_sf(data = pacifico, fill="#518FB5") +
  #geom_sf(data = ciudades_map, fill="#e34a33") +
  #coord_sf(xlim = bb[1, ], ylim = bb[2, ])
  theme_void()

```




Exploring ETC
====================================================
Sidebar {.sidebar}
-----------------------------------------------------------------------
**Guides for interpreting results** <br>
_Relevant years and areas_ <br>
The Colombian Institute for the Evaluation of Education (ICFES) applied different evaluations in different years. The first year of the tests - 2009 - was a pilot year where math, language, and natural science evaluations were applied to 5th and 9th grades. Between 2012 and 2017, the evaluations were applied every year for 3rd, 5th, and 9th grades in the areas of **mathematics** and **language**. In _even_ years, **natural sciences** were evaluated for 5th and 9th grades, and in _odd years_, **citizenship skills** were evaluated again for 5th and 9th grades.
<br>
_Average scores_ <br>
Average scores are reported for each ETC. Due to the sampling nature of the tests, there is a margin of error and uncertainty. It is appropriate to interpret the results with this margin in mind. <br> 
_Performance levels_ <br>
The ICFES categorizes performance levels into 4 categories: **advanced**, **satisfactory**, **minimum**, and **insufficient**. To learn more about the meaning of each of these categories, visit (link).


Row {data-height=75 data-orientation=columns, layout=scroll}
-----------------------------------------------------------------------

```{r}
### Widgets 
selectInput("location", 
            label = "Select location", 
            choices = etcs_filtradas,
            selected = "Cali")

selectInput("grade_level", 
            label = "Select grade level", 
            choices = grados,
            selected = "Grado3")
```

Row {data-height=320 data-orientation=columns, layout=scroll}
-----------------------------------------------------------------------
### Average historical scores - MATH

```{r}
renderPlot({
### Set values to filter for
#ciudad_temp <-  input$location #"Cartago" #location
#grado <-     input$grade_level #"Grado5"     #grade_level
#area <- "Matematicas"

########################################################################
################     DATA MANIPULATION                   ############### 
########################################################################
  
### Filter df for ETC, grade, subject area, and relevant columns
df_filered_temp <- df_filered[df_filered$ENTIDAD == input$location,]
df_filered_temp <- df_filered_temp[df_filered_temp$grado == input$grade_level,]
df_filered_temp <- df_filered_temp[df_filered_temp$area == "Matematicas",]
df_filered_temp <- df_filered_temp %>% select("ENTIDAD", "PUNTAJE_PROMEDIO", "DESVIACION", "N", "ano")

### Create values std
df_filered_temp$std_high <- df_filered_temp$PUNTAJE_PROMEDIO + df_filered_temp$DESVIACION
df_filered_temp$std_low <- df_filered_temp$PUNTAJE_PROMEDIO - df_filered_temp$DESVIACION


########################################################################
################     PLOTTING                            ############### 
########################################################################

# Create a line plot with std
plot_mate_punt <- ggplot(df_filered_temp, aes(x = ano, y = PUNTAJE_PROMEDIO)) +
  geom_line(color='#2c7bb6', size=1) + 
  geom_point(aes(x = ano, y = PUNTAJE_PROMEDIO), size=3, shape=18) +
  geom_ribbon(aes(ymin = std_low, ymax = std_high), alpha = 0.4, fill='#abd9e9' ) + ### produce error shade
  labs(
          x= "Año",
          y= "",
          subtitle = "[puntaje, banda representa una desviación estandar desde el promedio]",
          title = "Puntajes promedios en matemáticas para la ETC + ciudad_temp \n en el grado + grado",
          caption = "Source: ICFES"
          ) +
  scale_x_continuous(breaks=c(unique(df_filered_temp$ano))) +
  theme_minimal() + 
  theme(
          panel.grid.major.y = element_blank(), ### remove gridlines
          panel.grid.minor.y = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.x = element_blank(), ### element_line(size = 1.1),  ### keep major gridlines in x-axis
          plot.title = element_text(size=14),
          #aspect.ratio = .5
      ) 
plot_mate_punt

}) ## ends interactive plot

```


### Percentage of students at proficiency levels - MATH

```{r}
renderPlot({                         ### activates interactive plot
  
########################################################################
################     DATA MANIPULATION                   ############### 
########################################################################

### Set values to filter for --- from INPUTS! 
#ciudad_temp <- input$location
#grado <- input$grade_level
#area <- "Matematicas"

### Filter df for ETC, grade, subject area, and relevant columns
df_filered_temp <- df_filered[df_filered$ENTIDAD == input$location,]
df_filered_temp <- df_filered_temp[df_filered_temp$grado == input$grade_level,]
df_filered_temp <- df_filered_temp[df_filered_temp$area == "Matematicas",]
df_filered_temp <- df_filered_temp %>% select("ENTIDAD", "PORCENTAJE_INSUFICIENTE", "PORCENTAJE_MINIMO",        
                                                "PORCENTAJE_SATISFACTORIO", "PORCENTAJE_AVANZADO","N", "ano")

### Create pivot for easier visual
df_long <- pivot_longer(df_filered_temp, cols = starts_with("PORCENTAJE"), 
                        names_to = "nivel_desemp", values_to = "porcentaje")

### Recode performance levels 
df_long$nivel_desemp <- gsub(".*_", "", df_long$nivel_desemp)

### Specify order of levels
order_levels <- c("AVANZADO", "SATISFACTORIO", "MINIMO", "INSUFICIENTE")
df_long$ord_nivel_desemp <- factor(df_long$nivel_desemp, levels=order_levels)  ### creates new var to reorder!


########################################################################
################     PLOTTING                            ############### 
########################################################################

# Create a stacked area plot
plot_mate_desemp <- ggplot(df_long, aes(x = ano, y = porcentaje, fill = ord_nivel_desemp)) +
  geom_area(alpha=0.9) + 
  labs(
          x= "Año",
          y= "",
          subtitle = "[porcentaje]",
          title = "Niveles de desempeño en matemáticas para la ETC + ciudad_temp \n en el grado + grado",
          caption = "Source: ICFES",
          fill = "Nivel de desempeño"
          ) +
  scale_fill_manual(values = c('AVANZADO'='#2c7bb6', 'SATISFACTORIO'='#abd9e9', 
                               'MINIMO'='#ffffbf', 'INSUFICIENTE'='#fdae61')) + ### alternative greens (#a6d96a, #1a9641)
  scale_x_continuous(breaks=c(unique(df_long$ano))) +
  theme_minimal() + 
  theme(
          panel.grid.major.y = element_blank(), ### remove gridlines
          panel.grid.minor.y = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.x = element_line(size = 1.1),  ### keep major gridlines in x-axis
          plot.title = element_text(size=14),
          #aspect.ratio = .5
      ) 
  # guides(fill = guide(title = "Proficiency level", #adjust legend
  #                              title.position="left",
  #                              #label.position = "top",
  #                              orientation = "vertical"))

plot_mate_desemp

}) ## ends interactive plot

```


Row {data-height=320 data-orientation=columns}
-----------------------------------------------------------------------
### Average historical scores - READING

```{r}
renderPlot({                         ### activates interactive plot

########################################################################
################     DATA MANIPULATION                   ############### 
########################################################################

### Set values to filter for --- from INPUTS! 
#ciudad_temp <- input$location
#grado <- input$grade_level
#area <- "Lenguaje"


### Filter df for ETC, grade, subject area, and relevant columns
df_filered_temp <- df_filered[df_filered$ENTIDAD == input$location,]
df_filered_temp <- df_filered_temp[df_filered_temp$grado == input$grade_level,]
df_filered_temp <- df_filered_temp[df_filered_temp$area == "Lenguaje",]
df_filered_temp <- df_filered_temp %>% select("ENTIDAD", "PUNTAJE_PROMEDIO", "DESVIACION", "N", "ano")

### Create values std
df_filered_temp$std_high <- df_filered_temp$PUNTAJE_PROMEDIO + df_filered_temp$DESVIACION
df_filered_temp$std_low <- df_filered_temp$PUNTAJE_PROMEDIO - df_filered_temp$DESVIACION

########################################################################
################     PLOTTING                            ############### 
########################################################################

# Create a line plot with std
plot_lect_punt <- ggplot(df_filered_temp, aes(x = ano, y = PUNTAJE_PROMEDIO)) +
  geom_line(color='#2c7bb6', size=1) + 
  geom_point(aes(x = ano, y = PUNTAJE_PROMEDIO), size=3, shape=18) +
  geom_ribbon(aes(ymin = std_low, ymax = std_high), alpha = 0.4, fill='#abd9e9' ) + ### produce error shade
  labs(
          x= "Año",
          y= "",
          subtitle = "[puntaje, banda representa una desviación estandar desde el promedio]",
          title = "Puntajes promedios en lectura para la ETC + ciudad_temp \n en el grado + grado",
          caption = "Source: ICFES"
          ) +
  scale_x_continuous(breaks=c(unique(df_filered_temp$ano))) +
  theme_minimal() + 
  theme(
          panel.grid.major.y = element_blank(), ### remove gridlines
          panel.grid.minor.y = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.x = element_blank(), ### element_line(size = 1.1),  ### keep major gridlines in x-axis
          plot.title = element_text(size=14),
          #aspect.ratio = .5
      ) 
plot_lect_punt

}) ## ends interactive plot
```



### Percentage of students at proficiency levels - READING

```{r}
renderPlot({                         ### activates interactive plot

########################################################################
################     DATA MANIPULATION                   ############### 
########################################################################

### Set values to filter for --- from INPUTS! 
#ciudad_temp <- input$location
#grado <- input$grade_level
#area <- "Lenguaje"

### Filter df for ETC, grade, subject area, and relevant columns
df_filered_temp <- df_filered[df_filered$ENTIDAD == input$location,]
df_filered_temp <- df_filered_temp[df_filered_temp$grado == input$grade_level,]
df_filered_temp <- df_filered_temp[df_filered_temp$area == "Lenguaje",]
df_filered_temp <- df_filered_temp %>% select("ENTIDAD", "PORCENTAJE_INSUFICIENTE", "PORCENTAJE_MINIMO",        
                                                "PORCENTAJE_SATISFACTORIO", "PORCENTAJE_AVANZADO","N", "ano")

### Create pivot for easier visual
df_long <- pivot_longer(df_filered_temp, cols = starts_with("PORCENTAJE"), 
                        names_to = "nivel_desemp", values_to = "porcentaje")

### Recode performance levels 
df_long$nivel_desemp <- gsub(".*_", "", df_long$nivel_desemp)

### Specify order of levels
order_levels <- c("AVANZADO", "SATISFACTORIO", "MINIMO", "INSUFICIENTE")
df_long$ord_nivel_desemp <- factor(df_long$nivel_desemp, levels=order_levels)  ### creates new var to reorder!


########################################################################
################     PLOTTING                            ############### 
########################################################################

# Create a stacked area plot
plot_lect_desemp <- ggplot(df_long, aes(x = ano, y = porcentaje, fill = ord_nivel_desemp)) +
  geom_area(alpha=0.9) + 
  labs(
          x= "Año",
          y= "",
          subtitle = "[porcentaje]",
          title = "Niveles de desempeño en lectura para la ETC + ciudad_temp \n en el grado + grado",
          caption = "Source: ICFES",
          fill = "Nivel de desempeño"
          ) +
  scale_fill_manual(values = c('AVANZADO'='#2c7bb6', 'SATISFACTORIO'='#abd9e9', 
                               'MINIMO'='#ffffbf', 'INSUFICIENTE'='#fdae61')) + ### alternative greens (#a6d96a, #1a9641)
  scale_x_continuous(breaks=c(unique(df_long$ano))) +
  theme_minimal() + 
  theme(
          panel.grid.major.y = element_blank(), ### remove gridlines
          panel.grid.minor.y = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.x = element_line(size = 1.1),  ### keep major gridlines in x-axis
          plot.title = element_text(size=14),
          #aspect.ratio = .5
      ) 
  # guides(fill = guide(title = "Proficiency level", #adjust legend
  #                              title.position="left",
  #                              #label.position = "top",
  #                              orientation = "vertical"))

plot_lect_desemp

}) ## ends interactive plot
```


Row {data-height=320 .tabset}
-----------------------------------------------------------------------
### Natural Science

(some grade levels will display the assessments for natural science, though only for selected years)

```{r}

```

### Social Studies 

(some grade levels will display the assessments for social studies, though only for selected years)


```{r}

```


Comparing ETCs
====================================================

Sidebar {.sidebar}
-----------------------------------------------------------------------

```{r}
### TRIAL, removed below
##Row {data-height=25 data-orientation=columns, layout=scroll}  data-width=50

### Widgets 

checkboxGroupInput("checkbox", 
                   label = "Select locations",
                   choices = etcs_filtradas,
                   selected = c("Cali", "Popayán", "Pasto","Quibdó"))

### Removed, need to select multiple in checkbox above 
#selectInput("location_compared", label = "Select location", choices = etcs_filtradas)

selectInput("grade_level_compared", 
            label = "Select grade level", 
            choices = grados,
            selected = "Grado3")

selectInput("areas_compared", 
            label = "Select area", 
            choices = areas,
            selected = "Matematicas")

```

Row {data-height=600 data-orientation=columns, layout=scroll}
-----------------------------------------------------------------------
### Comparting ETCs

```{r}
renderPlot({                         ### activates interactive plot

########################################################################
################     DATA MANIPULATION                   ############### 
########################################################################

### Set values to filter for --- from INPUTS! 
#ciudad_temp <- input$location_compared
#grado <- input$grade_level_compared
#area <- input$areas_compared

### Filter df for ETC, grade, subject area, and relevant columns
df_filered_comp <- df_filered ###[df_filered$ENTIDAD %in% input$checkbox,], now map all and highlight 
df_filered_comp <- df_filered_comp[df_filered_comp$grado == input$grade_level_compared,]
df_filered_comp <- df_filered_comp[df_filered_comp$area == input$areas_compared,]
df_filered_comp <- df_filered_comp %>% select("ENTIDAD", "PUNTAJE_PROMEDIO", "DESVIACION", "N", "ano")

### Create values std
df_filered_comp$std_high <- df_filered_comp$PUNTAJE_PROMEDIO + df_filered_comp$DESVIACION
df_filered_comp$std_low <- df_filered_comp$PUNTAJE_PROMEDIO - df_filered_comp$DESVIACION

########################################################################
################     PLOTTING                            ############### 
########################################################################

# Create a line plot without stds
plot_comparaciones <- ggplot(df_filered_comp, aes(x = ano, y = PUNTAJE_PROMEDIO, color=ENTIDAD)) +
  geom_line(size=2) + 
  gghighlight(ENTIDAD %in% input$checkbox, 
              unhighlighted_params = list(linewidth = 1),
              label_params = list(size = 8)) +  ## line_label_type = "label_path",
  geom_point(aes(x = ano, y = PUNTAJE_PROMEDIO), size=5, shape=18) +
  #geom_ribbon(aes(ymin = std_low, ymax = std_high), alpha = 0.4, fill='#abd9e9' ) + ### produce error shade
  labs(
          x= "Año",
          y= "",
          #subtitle = "[puntaje, banda representa una desviación estandar desde el promedio]",
          title = "Puntajes promedios en (area) en (grado)",
          caption = "Source: ICFES"
          ) +
  scale_x_continuous(breaks=c(unique(df_filered_comp$ano))) +
  scale_color_viridis(discrete=TRUE) +
  theme_minimal() + 
  theme(
          panel.grid.major.y = element_blank(), ### remove gridlines
          panel.grid.minor.y = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.x = element_blank(), ### element_line(size = 1.1),  ### keep major gridlines in x-axis?
          plot.title = element_text(size=18),
          axis.text = element_text(size=14),
          axis.title.x = element_text(size=16),
          #aspect.ratio = .5
      ) 
plot_comparaciones

}) ## ends interactive plot
```

```{r eval=FALSE}
########################################################################
################     DATA MANIPULATION                   ############### 
########################################################################

### Set values to filter for --- from INPUTS! 
#ciudad_temp <- input$location_compared
#grado <- input$grade_level_compared
#area <- input$areas_compared

### Filter df for ETC, grade, subject area, and relevant columns
#df_filered_comp <- df_filered[df_filered$ENTIDAD %in% input$checkbox,]
#df_filered_comp <- df_filered_comp[df_filered_comp$grado == input$grade_level_compared,]
#df_filered_comp <- df_filered_comp[df_filered_comp$area == input$areas_compared,]
  
###TEMP
ciudades_temp <- c("Cali", "Yumbo", "Pasto","Palmira")
grado_temp <- "Grado5"
area_temp <- "Matematicas"
df_filered_comp <- df_filered##[df_filered$ENTIDAD %in% ciudades_temp,]
df_filered_comp <- df_filered_comp[df_filered_comp$grado == grado_temp,]
df_filered_comp <- df_filered_comp[df_filered_comp$area == area_temp,]
##END TEMP

df_filered_comp <- df_filered_comp %>% select("ENTIDAD", "PUNTAJE_PROMEDIO", "DESVIACION", "N", "ano")

## Reorder based on max value
max_year <- max(df_filered_comp$ano)
order <- df_filered_comp[df_filered_comp$ano == max_year,]
order$ENTIDAD <- reorder(order$ENTIDAD, order$PUNTAJE_PROMEDIO, FUN= max)
#df_filered_comp$ENTIDAD <- reorder(df_filered_comp$ENTIDAD, df_filered_comp$PUNTAJE_PROMEDIO, FUN= max)

### Create values std
#df_filered_comp$std_high <- df_filered_comp$PUNTAJE_PROMEDIO + df_filered_comp$DESVIACION
#df_filered_comp$std_low <- df_filered_comp$PUNTAJE_PROMEDIO - df_filered_comp$DESVIACION

########################################################################
################     PLOTTING                            ############### 
########################################################################

# Create a line plot with std
plot_comparaciones <- ggplot(df_filered_comp, aes(x = ano, y = PUNTAJE_PROMEDIO, color=ENTIDAD)) +
  geom_line(size=1) + 
  gghighlight(ENTIDAD %in% ciudades_temp, unhighlighted_params = list(linewidth = 1),
              line_label_type = "label_path", label_params = list(size = 10)) +  ##  
  geom_point(aes(x = ano, y = PUNTAJE_PROMEDIO), size=3, shape=18) +
  #geom_ribbon(aes(ymin = std_low, ymax = std_high), alpha = 0.4, fill='#abd9e9' ) + ### produce error shade
  labs(
          x = "Ano",
          y = " ", 
          title = "Puntajes promedios en {area_temp} en {grado_temp}",
          caption = "Source: ICFES",
          ) +
  scale_x_continuous(breaks=c(unique(df_filered_comp$ano))) +
  scale_color_viridis(discrete=TRUE) +
  theme_minimal() + 
  theme(
          panel.grid.major.y = element_blank(), ### remove gridlines
          panel.grid.minor.y = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.x = element_blank(), ### element_line(size = 1.1),  ### keep major gridlines in x-axis
          plot.title = element_text(size=18),
          axis.text = element_text(size=16),
          axis.title.x = element_text(size=14),
          #aspect.ratio = .5
      ) 
  
plot_comparaciones
```


View code
=======================================================================

```{r eval=FALSE}
 # Add a button to view the source code
 actionButton("view_code", "View Source Code")

 # Render the code when the button is clicked
 renderPrint({
   # Get the source code of the dashboard
   code <- capture.output({
     cat("```{r}\n")
     cat(readLines("my_dashboard.Rmd"), sep = "\n")
     cat("\n```")
   })

   # Display the code
   code
 })
```


Translate
=======================================================================

```{r eval=FALSE}
# Add a button to translate to English
actionButton("translate", "Translate to English")

# Define a reactive function to translate the text
translated_text <- reactive({
  if (input$translate) {
    # Translate the text to English using googleLanguageR
    translate_text(text = "El contenido de mi dashboard en español", target = "en")
  } else {
    # Return the original text
    "El contenido de mi dashboard en español"
  }
})

# Display the translated text
renderText({
  translated_text()
})
```







